%==============================================================================
\chapter{Budget envelope and resources}
%==============================================================================

\section{Framework}
The project falls within an \textbf{academic (mini-project)} framework. The
budget envelope therefore mainly translates into the \textbf{human, software and
hardware resources} mobilised, with no direct financial cost.

\section{Mobilised resources}
\begin{xltabular}{\textwidth}{>{\bfseries}p{4cm} X}
\toprule
\textbf{Resource} & \textbf{Detail} \\
\midrule
\endhead
Human    & Student project team (design, development, testing, documentation). \\
Software & Containerised runtime environment, relational DBMS, IDE, UML tools, compilation chain. \\
Hardware & Development workstations, simulated sensors for the indicators part. \\
External & Google OAuth account, X.509 certificates (self-signed in the demonstration). \\
\bottomrule
\end{xltabular}

\begin{encadre}
\textbf{Major academic constraint:} any use of code generators (ChatGPT,
DeepSeek and derivatives) is prohibited and will be considered fraud in the
examination.
\end{encadre}

%------------------------------------------------------------------------------
\section{Economic valuation: industrialisation hypothesis}
%------------------------------------------------------------------------------
The previous section describes the \emph{actual} cost of the project, nil in
direct spending. As a valuation exercise, this section estimates what the same
product would cost \textbf{if delivered by a software services company}. The
amounts are \textbf{working hypotheses}, not committed expenditure.

\subsection{Reference workload}
The workload plan in the schedule chapter retains \textbf{304 story points}
across 8~sprints. At $0.79$~person-days per point, the workload amounts to
$\approx 240$~person-days; at $1$~person-day per point --- a prudent assumption
for a team new to the stack --- it reaches $\approx 304$~person-days. This range
forms the basis of the costing.

\subsection{Breakdown of a reference envelope}
For an envelope of \textbf{USD 200,000}, the following breakdown reflects market
practice. Development is not the majority share: equating it with the total cost
is a classic costing mistake.

\begin{xltabular}{\textwidth}{>{\bfseries}p{5.2cm} r r X}
\toprule
\textbf{Item} & \textbf{\%} & \textbf{USD} & \textbf{Rationale} \\
\midrule
\endhead
Development & 45 & 90,000 & Back-end, front-end, integration \\
Project management / analysis & 12 & 24,000 & The most frequently underestimated item \\
Quality and testing & 10 & 20,000 & Coverage $\geq 70\,\%$, integration and load testing \\
UX/UI design and accessibility & 8 & 16,000 & Reduced cost: the design system already exists \\
Infrastructure and tooling (year 1) & 6 & 12,000 & Hosting, backups, monitoring \\
External security audit & 5 & 10,000 & Sensitive geolocation data \\
Contingency reserve & 14 & 28,000 & Without a reserve, an envelope of this size overruns \\
\bottomrule
\end{xltabular}

\subsection{Sensitivity to the daily rate}
The dominant factor is geographic rather than technical. At an average rate of
USD 600--900 per person-day, the envelope funds $\approx 220$--$280$~person-days:
it barely covers the reference workload, \textbf{with no margin}. At a rate of
USD 250--400, it funds $\approx 500$--$700$~person-days and allows a team of five
to six people together with a sound contingency reserve.

\subsection{Adjustable scope}
Should the envelope tighten, the following order of removal preserves the core of
the product: first the advanced features (annex: queen tracking, batch QR code),
then adaptive anomaly detection --- whose calibration without real historical
data is an identified risk --- and finally live sensor ingestion, the data model
being sufficient for a first version. The business entities, visits,
authentication and access control, offline mode and dashboards remain
incompressible.

\subsection{Total cost of ownership}
Build cost represents only part of the product's lifetime cost. Corrective and
evolutionary maintenance is usually costed at \textbf{15--20\,\% per year} of the
build envelope, i.e.\ USD 30,000--40,000 annually. Over three years, the total
cost of ownership therefore approaches \textbf{USD 300,000}. To this must be
added the items regularly omitted from an initial costing: migration of the
operator's existing data, training and change management for field users,
professionally reviewed translation for the three languages, and compliance work
relating to personal data.

\begin{encadre}
\textbf{Scope of this section:} this is an educational \emph{valuation} intended
to situate the project's effort within market orders of magnitude. It commits no
expenditure and does not alter the academic framework defined at the beginning of
this chapter.
\end{encadre}
