\chapter{Glossary}
%==============================================================================
\begin{xltabular}{\textwidth}{>{\bfseries}p{4cm} X}
\toprule
\textbf{Term} & \textbf{Definition} \\
\midrule
\endhead
Base or body & Base compartment of the hive holding the brood frames. \\
Super & Extension level added to the body for honey harvesting. \\
Frame & Removable support of the wax combs, installed on a slot. \\
Slot & Identified location of a frame in a body or a super. \\
Swarm & Colony of bees populating a hive. \\
Swarming & Departure of part of the colony with the queen to found a new colony. \\
Nucleus & Small colony formed to split a swarm or raise a queen. \\
Queen & The colony's single reproductive female. \\
Brood & All the eggs, larvae and pupae of the colony. \\
Foraging & Activity of collecting nectar and pollen by the bees. \\
Propolis & Resinous substance collected by the bees. \\
Queen excluder & Grid that confines the queen to the body and separates the supers. \\
Feeder & Device for the food replenishment of the colony. \\
Configurable threshold & Configurable reference value triggering an alert. \\
ROI & Return on investment, ratio between valued production and costs. \\
CRUD & Create, read, update and delete operations. \\
i18n & Internationalisation, adaptation to languages and unit systems. \\
Spring Boot & Java application framework: dependency injection, autoconfiguration and embedded server. \\
JPA & Standard mapping interface between Java objects and relational tables. \\
jOOQ & Library for typed SQL queries, checked at compile time. \\
OpenAPI & Contract describing a REST API, usable by humans and machines. \\
OAuth / OIDC & Protocols for delegated authorization and authentication. \\
X.509 & Certificates ensuring authentication of the parties in an encrypted exchange. \\
TLS & Encryption protocol for network exchanges (transport layer), basis of HTTPS. \\
REST / JSON & Web-API style and text format for exchanging data (measurement ingestion). \\
MQTT & Lightweight messaging protocol (publish/subscribe) suited to sensors and constrained networks. \\
JWT (token) & Signed identity-bearing token, issued by the OAuth provider. \\
RBAC & Role-based access control, applied server-side (see appendix~\ref{chap:complements}). \\
GDPR & General Data Protection Regulation. Governs personal data. \\
Retention & Duration for which a datum is kept before archiving, anonymisation or purge. \\
Encryption at rest & Encryption of stored data (sensitive columns) in the database. \\
Pseudonymisation & Dissociation of a person's identity from their data, reversible under control. \\
Idempotency & Property of an operation that can be replayed without creating a duplicate (measurement ingestion). \\
Hysteresis & Persistence of a breach over several measurements before raising an alert (debounce). \\
\bottomrule
\end{xltabular}
