%==============================================================================
\chapter{Schedule and deliverables}
\label{chap:delais}
%==============================================================================

\section{Expected deliverables}
\begin{itemize}
  \item Functional application (Java server source code + front-end).
  \item Test plan and execution results.
  \item Design file (UML diagrams).
  \item Project report, poster and oral presentation.
\end{itemize}

\section{Acceptance criteria}
Acceptance is pronounced when \textbf{all} the following criteria are satisfied
and jointly verified. Every criterion is measurable: a criterion that cannot be
measured is not an acceptance criterion.

\begin{xltabular}{\textwidth}{>{\bfseries}p{3.2cm} X p{3.4cm}}
\toprule
\textbf{Area} & \textbf{Criterion} & \textbf{Means of proof} \\
\midrule
\endhead
Functional & The use cases of chapter~\ref{chap:cas-utilisation} are implemented and demonstrated on real data. & Acceptance report \\
\addlinespace
Code quality & Business-layer test coverage $\geq$ 70~\%, no untreated critical or high vulnerability. & JaCoCo report, dependency scan \\
\addlinespace
Security & External penetration test carried out; no critical or high vulnerability open at delivery. & Pentest report and remediation plan \\
\addlinespace
Data protection & Geolocation coordinates inaccessible outside an authorised role, absent from default exports and from logs. & Authorisation tests, log review \\
\addlinespace
Performance & p95 $<$ 500~ms and error rate $<$ 1~\% under the nominal load defined in appendix~\ref{chap:exploitation}. & k6 load-test report \\
\addlinespace
Availability & The service (SLO), recovery (RTO) and maximum loss (RPO) objectives of appendix~\ref{chap:exploitation} are met. & Monitoring, restoration test \\
\addlinespace
Backup & A full restoration has been carried out successfully under real conditions. & Restoration test report \\
\addlinespace
Accessibility and usability & Interface operable by keyboard, contrasts compliant with the design system, journey validated by a business user. & Accessibility audit, user acceptance \\
\addlinespace
Internationalisation & The three languages (FR, EN, AR) are complete, Arabic RTL rendering verified on every screen. & Linguistic review \\
\addlinespace
Operability & Runbook delivered, knowledge transfer completed, system redeployable from the repository. & Transfer session, dry-run redeployment \\
\addlinespace
Reversibility & Data exportable in an open, documented format; source code and documentation handed over. & Demonstration export \\
\bottomrule
\end{xltabular}

\begin{encadre}
\textbf{Reservations.} A minor reservation does not prevent acceptance but is
subject to an agreed clearance deadline. A \textbf{blocking} reservation ---
any critical vulnerability, any restoration failure, any data leak between
holdings --- suspends acceptance until correction and re-verification.
\end{encadre}

\section{UML diagrams to produce}
Diagrams of: use case, classes, objects, sequence, activity, state machine,
interaction / collaboration, package, component and deployment. The expected
content of each is detailed in chapter \ref{chap:conception} on the design.

\section{Provisional schedule}
The schedule is expressed in relative weeks, with the intermediate deliverables
associated with each phase.
\begin{xltabular}{\textwidth}{>{\bfseries}p{4.5cm} C{2.2cm} X}
\toprule
\textbf{Phase} & \textbf{Weeks} & \textbf{Deliverable} \\
\midrule
\endhead
Analysis and requirements specification & W1 to W2 & Validated requirements specification \\
Design (UML) & W3 to W4 & Design file \\
CRUD core development & W5 to W7 & Model and entity management \\
Dashboards and alerts & W8 to W9 & Calendar, alerts and summaries \\
Web services, security, i18n & W10 to W11 & OAuth, API, X.509 and SSL, XML \\
Testing and validation & W12 & Executed test plan \\
Documentation and presentation & W13 & Report, poster and defence \\
\bottomrule
\end{xltabular}
